Variational data assimilation for the wave equation in heterogeneous media

### 2025-06-27 tex/main.tex
update refs to homogeneous paper and minors

\affil[2]{\orgdiv{MAKUTU}, \orgname{Inria}, \orgaddress{\street{...}, \city{...}, \postcode{...}, \country{France}}}

@@

\section{Introduction}
\todo[inline]{@JP: please update affiliation}

@@

\begin{proof}
    It suffices to show that $\tnorm{(\cdot,\cdot)}$ is a norm. Since $\tnorm{(\cdot,\cdot)}$ is a semi-norm by definition, we only require that $\tnorm{(\Uh,\Zh)} = 0$ implies $\Uh=\Zh= 0$. Assume that $\tnorm{(\Uh,\Zh)} = 0$. Then, in particular, $\vert \Uh \vert_{\uparrow \downarrow} = 0$ and therefore $(\ul_1,\ul_2) \in [H^1(Q)]^2$. Furthermore, by definition of the stabilization terms $S_h$ and $S_h^{\ast}$, we have that $\Zh = 0$, $\ul_1 \vert_{\Sigma} = 0$, $\ul_1 \vert_{\omega_T} = 0$ and $\dt \ul_1 = \ul_2$. Similar to \cite[Lem. 2.1]{BP24}, partial integration yields 
    \begin{align*}
	   & \Vert \wop \ul_1 \Vert^2_{H^{-1}(Q)} := \sup_{\substack{  y \in H^1_0(Q), \\ \norm{y}_{H^1(Q) } = 1  }} \int_{Q} \left\{ -(\dt \ul_1) \dt y + c^2 \nabla \ul_1 \nabla y \right\}  \\
        &= \sup_{\substack{  y \in H^1_0(Q), \\ \norm{y}_{H^1(Q) } = 1  }} \Big\{ \sum_{n = 0}^{N-1} (\ul_2 - \dt \ul_1, \dt y)_{Q_n} 
        + \sum_{n = 0}^{N-1} \int_{I_n} \sum_{K \in \mathcal{T}_h} (\dt \ul_2 - \div(c^2 \nabla \ul_1),y)_{K} \ \dT \\
	    & \hspace{6em} + \sum_{n = 0}^{N-1} \int_{I_n} \sum_{F \in \mathcal{F}_i} (\jump{c^2 \nabla \ul_1} \cdot \mathbf{n}, y)_F \ \dT + \sum_{n = 0}^{N-1} (\jump{\ul_1^n},y)_{\Sigma^n} \Big\},
    \end{align*}
where $\jump{c^2 \nabla \ul_1} \cdot \mathbf{n}$  denotes the jump of the normal derivative over interior facets. It follows by definition of the stabilizers and $\tnorm{(\Uh,\Zh)} = 0$ that $\Vert \wop \ul_1 \Vert_{H^{-1}(Q)} = 0$ and that the interface conditions \eqref{eq:IF-conditions} are fulfilled. Thus, by taking $\mu \rightarrow \infty$ in \cref{thm:filippas}, we conclude that $u_0 = u_1 = 0$. As solutions to \eqref{eq:waveEquation}-\eqref{eq:IF-conditions}+\eqref{eq:initialData} depend continuously on the data \cite{StolkPhD}, it follows that $\Uh = 0$. 
\end{proof}

@@

\begin{proof}
    The proof follows with the same arguments as the proof of \cite[Theorem 10]{BP24}.
\end{proof}

\begin{rmk}[Homogenous media]\label{rem:homMedia}
    If $c_i = 1$ for $i = 1,2$, \cref{assum:LipschitzStability} holds (see \cref{thm:Lipschitz}), and we recover the result from \cite[Theorem 10]{BP24}.
\end{rmk}

\begin{rmk}[Logarithmic stability]
    If \cref{assum:LipschitzStability} is not satisfied, we could still hope to utilize the results from \cref{thm:filippas}, provided that $T$ is large enough such that \eqref{eq:Tcondition} holds. To this end, we would have to deal with the following technicalities:  
    \begin{enumerate}
        \item The results from \cref{thm:filippas} are formulated in terms of $\Vert \wop u \Vert_{L^2(Q)}$, but we can only control $\Vert \wop \tilde{e}_h \Vert_{H^{-1}(Q)}$ for the residual $\tilde{e}_h = u - L_{\Delta t} \ul_1$. Thus, we would have to rewrite the estimates from \cref{thm:filippas} with $\Vert \wop u \Vert_{H^{-1}(Q)}$ instead of $\Vert \wop u \Vert_{L^2(Q)}$. To this end, one might be able to shift the regularity suitably, cf. \cite[Sec. 2.6]{StolkPhD}, or use similar arguments as in \cite[Thm. A.1]{BFMO21}.
        \item We have to ensure that the constants in the estimates from \cref{thm:filippas} remain bounded when we consider the residual $\tilde{e}_h$. In particular, we have to control $C_{0,1} := \Vert (\tilde{e}_h \vert_{t = 0}, \dt \tilde{e}_h \vert_{t = 0}) \Vert_{H^1 \times L^2}$. We note that we can always ensure that $C_{0,1}$ is bounded by adding a suitable Tikhonov regularization term to \eqref{eq:primalStab}. 
    \end{enumerate}
    Then, we can use the arguments of \cite[Theorem 10]{BP24} to show that 
    \begin{equation*}
        \Vert (u - \ul_1) \vert_{t = 0} \Vert_{L^2(\Omega)}  + \Vert \dt (u - \ul_1) \vert_{t = 0} \Vert_{H^{-1}(\Omega)} \le C C_{0,1} \log(1 + C_{0,1} h^{-s} \Vert u \Vert_{H^m(Q)}^{-1})^{-1}. 
    \end{equation*}
\end{rmk}

>>>>

 \affil[2]{\orgdiv{Inria Project-Team Makutu}, \orgname{Université de Pau et des Pays de l’Adour}, \orgaddress{\street{Avenue de l'Université - BP 1155}, \city{Pau}, \postcode{64013}, \country{France}}}

@@

\section{Introduction}

@@

\begin{proof}
    It suffices to show that $\tnorm{(\cdot,\cdot)}$ is a norm. Since $\tnorm{(\cdot,\cdot)}$ is a semi-norm by definition, we only require that $\tnorm{(\Uh,\Zh)} = 0$ implies $\Uh=\Zh= 0$. Assume that $\tnorm{(\Uh,\Zh)} = 0$. Then, in particular, $\vert \Uh \vert_{\uparrow \downarrow} = 0$ and therefore $(\ul_1,\ul_2) \in [H^1(Q)]^2$. Furthermore, by definition of the stabilization terms $S_h$ and $S_h^{\ast}$, we have that $\Zh = 0$, $\ul_1 \vert_{\Sigma} = 0$, $\ul_1 \vert_{\omega_T} = 0$ and $\dt \ul_1 = \ul_2$. Similar to \cite[Lemma 2]{BP24}, partial integration yields 
    \begin{align*}
	   & \Vert \wop \ul_1 \Vert^2_{H^{-1}(Q)} := \sup_{\substack{  y \in H^1_0(Q), \\ \norm{y}_{H^1(Q) } = 1  }} \int_{Q} \left\{ -(\dt \ul_1) \dt y + c^2 \nabla \ul_1 \nabla y \right\}  \\
        &= \sup_{\substack{  y \in H^1_0(Q), \\ \norm{y}_{H^1(Q) } = 1  }} \Big\{ \sum_{n = 0}^{N-1} (\ul_2 - \dt \ul_1, \dt y)_{Q_n} 
        + \sum_{n = 0}^{N-1} \int_{I_n} \sum_{K \in \mathcal{T}_h} (\dt \ul_2 - \div(c^2 \nabla \ul_1),y)_{K} \ \dT \\
	    & \hspace{6em} + \sum_{n = 0}^{N-1} \int_{I_n} \sum_{F \in \mathcal{F}_i} (\jump{c^2 \nabla \ul_1} \cdot \mathbf{n}, y)_F \ \dT + \sum_{n = 0}^{N-1} (\jump{\ul_1^n},y)_{\Sigma^n} \Big\},
    \end{align*}
where $\jump{c^2 \nabla \ul_1} \cdot \mathbf{n}$  denotes the jump of the normal derivative over interior facets. It follows by definition of the stabilizers and $\tnorm{(\Uh,\Zh)} = 0$ that $\Vert \wop \ul_1 \Vert_{H^{-1}(Q)} = 0$ and that the interface conditions \eqref{eq:IF-conditions} are fulfilled. Thus, by taking $\mu \rightarrow \infty$ in \cref{thm:filippas}, we conclude that $u_0 = u_1 = 0$. As solutions to \eqref{eq:waveEquation}-\eqref{eq:IF-conditions}+\eqref{eq:initialData} depend continuously on the data \cite{StolkPhD}, it follows that $\Uh = 0$. 
\end{proof}

@@

\begin{proof}
    The proof follows with the same arguments as the proof of \cite[Theorem 14]{BP24}.
\end{proof}

\begin{rmk}[Homogenous media]\label{rem:homMedia}
    If $c_i = 1$ for $i = 1,2$, \cref{assum:LipschitzStability} holds (see \cref{thm:Lipschitz}), and we recover the result from \cite[Theorem 14]{BP24}.
\end{rmk}

\begin{rmk}[Logarithmic stability]
    If \cref{assum:LipschitzStability} is not satisfied, we could still hope to utilize the results from \cref{thm:filippas}, provided that $T$ is large enough such that \eqref{eq:Tcondition} holds. To this end, we would have to deal with the following technicalities:  
    \begin{enumerate}
        \item The results from \cref{thm:filippas} are formulated in terms of $\Vert \wop u \Vert_{L^2(Q)}$, but we can only control $\Vert \wop \tilde{e}_h \Vert_{H^{-1}(Q)}$ for the residual $\tilde{e}_h = u - L_{\Delta t} \ul_1$. Thus, we would have to rewrite the estimates from \cref{thm:filippas} with $\Vert \wop u \Vert_{H^{-1}(Q)}$ instead of $\Vert \wop u \Vert_{L^2(Q)}$. To this end, one might be able to shift the regularity suitably, cf. \cite[Sec. 2.6]{StolkPhD}, or use similar arguments as in \cite[Thm. A.1]{BFMO21}.
        \item We have to ensure that the constants in the estimates from \cref{thm:filippas} remain bounded when we consider the residual $\tilde{e}_h$. In particular, we have to control $C_{0,1} := \Vert (\tilde{e}_h \vert_{t = 0}, \dt \tilde{e}_h \vert_{t = 0}) \Vert_{H^1 \times L^2}$. We note that we can always ensure that $C_{0,1}$ is bounded by adding a suitable Tikhonov regularization term to \eqref{eq:primalStab}. 
    \end{enumerate}
    Then, we can use the arguments of \cite[Theorem 14]{BP24} to show that 
    \begin{equation*}
        \Vert (u - \ul_1) \vert_{t = 0} \Vert_{L^2(\Omega)}  + \Vert \dt (u - \ul_1) \vert_{t = 0} \Vert_{H^{-1}(\Omega)} \le C C_{0,1} \log(1 + C_{0,1} h^{-s} \Vert u \Vert_{H^m(Q)}^{-1})^{-1}. 
    \end{equation*}
\end{rmk}
